### 2019-05-30 communication/explore.Rmd
analysis cleanup 2.

```{r}
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'thesis/code')
```

@@

```{r}
require(tidyverse)
require(meta)
```

@@

```{r}
#Scaled RR over time with initial RR > 1
madata %>% filter(outcome.measure == "Risk Ratio") %>% filter(!is.na(study.year)) %>% 
  group_by(file.nr, comparison.nr, subgroup.nr) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>% 
  mutate(first.effect = effect[which.min(study.year)]) %>% 
  filter(first.effect > 1) %>% 
  mutate(scaled.effect = scale(effect), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .25, alpha = 0.025)  + 
  geom_smooth(method = lm) + theme_bw() + 
  labs(title = "Scaled and centered RR over time", subtitle = "Initial finding > 1") + ylab("Risk Ratio")  + xlab("Time") + xlim(c(-4,4)) + ylim(c(-5,7))
```

@@

@@

@@

```{r}
madata %>% filter(!is.na(fishersz)) %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  ungroup %>% arrange(desc(n)) %>% mutate(trials = cumsum(n * nn)) %>% 
  filter(n < 50) %>% 
  full_join( madata %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>% 
               filter(n > 49) %>% ungroup %>% summarise(n = 50, nn = sum(nn))) %>%  
  ggplot(aes(x = n, y = trials)) + geom_point(col = "gray15") + 
  theme_bw() + labs(title = "Sum of Trials with Number of Replicate (Trials) >= n") + ylab("Number of trials") + xlab("n, number of replicates")
```

>>>>

```{r}
PATH_HOME = path.expand("~") # user home
PATH = file.path(PATH_HOME, 'thesis/cochrane')
```

@@

```{r}
# load(file = "cochrane.RData")

require(tidyverse)
require(meta)
```

@@

```{r}
#Scaled RR over time with initial RR > 1
madata %>% filter(outcome.measure == "Risk Ratio") %>% filter(!is.na(study.year)) %>% 
  group_by(file.nr, comparison.nr, subgroup.nr) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>% 
  mutate(first.effect = effect[which.min(study.year)]) %>% 
  filter(first.effect > 1) %>% 
  mutate(scaled.effect = scale(effect), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .25, alpha = 0.025)  + 
  geom_smooth(method = lm) + theme_bw() + 
  labs(title = "Scaled and centered RR over time", subtitle = "Initial finding > 1") + ylab("Risk Ratio")  + ylab("Time") + xlim(c(-4,4)) + ylim(c(-5,7))
```

@@

By hand computation of Cochrane's Q is not equal to Q computed with metabin package, as can be seen when comparing the histogramm. The shape of the histogramms suggest that the by hand computed one ist wrong (large right tail). The formula used for computing is:

\begin{align} 
Q = sum_{i=1}^{n} w_{i} (\hat{\theta} - \theta_{i}) 
\end{align}

with $n$ = number of trials, $w_{i}$ the inverse of the variance of the log effect estimate, $\theta_{i}$ the log effect estimate and $\hat{\theta}$ the weigthed average of the log effect estimates. The variance of the log odds ratio was computed as $(1/a + 1/b + 1/c + 1/d)$, $a,b,c,d$ being the entries in a two-by-two table.

Cochrane heterogeneity test for Odds Ratios (``manual''):

```{r}
madata %>% group_by(file.nr, outcome.nr, comparison.nr, subgroup.nr) %>% 
  filter(outcome.measure == "Odds Ratio") %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>%
  mutate(log.OR = log(effect), 
         se.log.OR = sqrt( (1/events1) + (1/(total1 - events1)) + (1/events2) + (1/(total2 - events2)))) %>% 
  filter(!is.na(se.log.OR) & se.log.OR != Inf) %>% mutate(summary.effect = sum( log.OR/se.log.OR^2 ) / sum( 1/se.log.OR^2 )) %>% 
  summarise(Q = sum(((log.OR - summary.effect)^2)/se.log.OR^2), n = n(), OR =  exp(min(summary.effect))) %>% 
  mutate(p.value = 1 - pchisq(Q, df = n - 1)) %>% 
  ggplot(aes(x = p.value)) + geom_histogram(col = "gray15", fill = "dodgerblue") +
  theme_bw() + labs(title = "Cochrane Heterogeneity Test P-values of ORs")
```

@@

For comparison, the single Q's are shown computed by the two methods:

```{r}
file.number.to.compare <- 23
ds <- madata %>% group_by(file.nr, outcome.nr, comparison.nr, subgroup.nr) %>% filter(file.nr == file.number.to.compare) %>% 
  filter(outcome.measure == "Odds Ratio") %>% filter(events1 > 0 & events2 > 0) %>% 
  mutate(counts = n()) %>% filter(counts > 1) %>%
  summarise(
    Q = metabin(event.e = events1, n.e = total1, event.c = events2, n.c = total2, studlab = study.name, method = "Inverse", sm = "OR")$Q,
    p.value = metabin(event.e = events1, n.e = total1, event.c = events2, n.c = total2, studlab = study.name, 
                      method = "Inverse", sm =  "OR")$pval.Q)

ds2 <- madata %>% group_by(file.nr, outcome.nr, comparison.nr, subgroup.nr) %>% filter(file.nr == file.number.to.compare) %>% 
  filter(outcome.measure == "Odds Ratio") %>% filter(events1 > 0 & events2 > 0) %>%
  mutate(counts = n()) %>% filter(counts > 1) %>%
  mutate(log.OR = log(effect), 
         se.log.OR = sqrt( (1/events1) + (1/(total1 - events1)) + (1/events2) + (1/(total2 - events2)))) %>% 
  filter(!is.na(se.log.OR)) %>% mutate(summary.effect = sum( log.OR/se.log.OR^2 ) / sum( 1/se.log.OR^2 )) %>% 
  summarise(Q = sum(((log.OR - summary.effect)^2)/se.log.OR^2), n = n(), OR =  exp(min(summary.effect))) %>% 
  mutate(p.value = 1 - pchisq(Q, df = n - 1))

identical(ds$Q, ds2$Q)
print(data_frame(metabin = ds$Q, manual = ds2$Q, ds$p.value, ds2$p.value), n = 11)
```

One example where the Qs were different:

```{r}
madata %>% filter(file.nr == 23, outcome.nr == 3, comparison.nr == 1, subgroup.nr == 3) %>% mutate(log.OR = log(effect), 
         se.log.OR = sqrt( (1/events1) + (1/(total1 - events1)) + (1/events2) + (1/(total2 - events2)))) %>% 
  filter(!is.na(se.log.OR) ) %>% mutate(summary.effect = sum( log.OR/se.log.OR^2 ) / sum( 1/se.log.OR^2 )) %>% 
  select(effect, events1, events2, total1, total2, log.OR, se.log.OR, summary.effect)
```

```{r}

madata %>% filter(file.nr == 23, outcome.nr == 3, comparison.nr == 1, subgroup.nr == 3) %>%
mutate(log.OR = log(effect), 
         se.log.OR = sqrt( (1/events1) + (1/(total1 - events1)) + (1/events2) + (1/(total2 - events2)))) %>% 
  filter(!is.na(se.log.OR)) %>% mutate(summary.effect = sum( log.OR/se.log.OR^2 ) / sum( 1/se.log.OR^2 )) %>% 
  summarise(Q = sum(((log.OR - summary.effect)^2)/se.log.OR^2), n = n())
```

Cochrane heterogeneity test for Risk Ratios:
```{r}
# madata %>% group_by(file.nr, outcome.nr, comparison.nr, subgroup.nr) %>% 
#   filter(outcome.measure == "Risk Ratio") %>% 
#   mutate(counts = n()) %>% filter(counts > 1) %>%
#   mutate(log.RR = log(effect), 
#          se.log.RR = sqrt( (total1 - events1)/(events1*total1)) + (total2 - events2)/(events2*total2)) %>% 
#   filter(!is.na(se.log.RR)) %>% mutate(summary.effect = sum( log.RR/se.log.RR^2 ) / sum( 1/se.log.RR^2 )) %>% 
#   summarise(Q = sum(((log.RR - summary.effect)^2)/se.log.RR^2), n = n(), RR =  exp(min(summary.effect))) %>% 
#   mutate(p.value = 1 - pchisq(Q, df = n - 1)) %>% 
#   ggplot(aes(x = p.value)) + geom_histogram(col = "gray15", fill = "dodgerblue") +
#   theme_bw() + labs(title = "Cochrane Heterogeneity Test P-values of RRs")
```

@@

```{r}
madata %>% filter(!is.na(fishersz)) %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>%
  ungroup %>% arrange(desc(n)) %>% mutate(trials = cumsum(n * nn)) %>% 
  filter(n < 50) %>% 
  full_join( madata %>% group_by(file.nr, comparison.nr, outcome.nr, subgroup.nr) %>% count %>% group_by(n) %>% count %>% 
               filter(n > 49) %>% ungroup %>% summarise(n = 50, nn = sum(nn))) %>%  
  ggplot(aes(x = n, y = trials)) + geom_point(col = "gray15") + 
  theme_bw() + labs(title = "Total Number of Trials with Number of Replicates >= n") + ylab("Number of trials") + xlab("n, number of replicates")
```

### 2019-05-30 report/chapter02.Rnw
analysis cleanup 2.

file.bin <- "pb.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else {
  meta.bin <- meta.bin.complete(data.ext, min.study.number = 10, sig.level = 0.05, sm = "OR")
  save(meta.bin, file =  file.path(PATH_RESULTS, file.bin))
}

file.cont <- "pb.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else {
  meta.cont <- meta.cont.complete(data.ext, min.study.number = 10, sig.level = 0.05)
  save(meta.cont, file =  file.path(PATH_RESULTS, file.cont))
}

file.meta <- "meta.RData"
if (file.exists(file.path(PATH_RESULTS, file.meta))) {
  load(file.path(PATH_RESULTS, file.meta))
} else {
  meta <- pb.meta.merge(meta.bin, meta.cont)
  save(meta, file =  file.path(PATH_RESULTS, file.meta))
}

file.cont <- "mly.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else {
  data.cont <- mly.cont(data.ext, 0.05, min.study.number = 2)
  save(data.cont, file =  file.path(PATH_RESULTS, file.cont))
}

file.bin <- "mly.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else {
  data.bin <- mly.bin(data.ext, 0.05, min.study.number = 2)
  save(data.bin, file =  file.path(PATH_RESULTS, file.bin))
}

load(file.path(PATH_RESULTS, file = "mly.RData"))
load(file.path(PATH_RESULTS, file = "data.processed.RData"))

require(biostatUZH)
require(tidyverse)
require(meta)
require(metasens)
require(gridExtra)

>>>>

load(file.path(PATH_RESULTS, "meta_analyses_summary_complete.RData"))
load(file.path(PATH_RESULTS, "meta.bin.RData"))
load(file.path(PATH_RESULTS, "meta.cont.RData"))
load(file.path(PATH_RESULTS, "meta.surv.RData"))
load(file.path(PATH_RESULTS, "data_used_for_analysis.RData"))

### 2019-05-30 report/chapter03.Rnw
analysis cleanup 2.

file.bin <- "pb.bin.RData"
if (file.exists(file.path(PATH_RESULTS, file.bin))) {
  load(file.path(PATH_RESULTS, file.bin))
} else {
  meta.bin <- meta.bin.complete(data.ext, min.study.number = 10, sig.level = 0.05, sm = "OR")
  save(meta.bin, file =  file.path(PATH_RESULTS, file.bin))
}

file.cont <- "pb.cont.RData"
if (file.exists(file.path(PATH_RESULTS, file.cont))) {
  load(file.path(PATH_RESULTS, file.cont))
} else {
  meta.cont <- meta.cont.complete(data.ext, min.study.number = 10, sig.level = 0.05)
  save(meta.cont, file =  file.path(PATH_RESULTS, file.cont))
}

file.meta <- "meta.RData"
if (file.exists(file.path(PATH_RESULTS, file.meta))) {
  load(file.path(PATH_RESULTS, file.meta))
} else {
  meta <- pb.meta.merge(meta.bin, meta.cont)
  save(meta, file =  file.path(PATH_RESULTS, file.meta))
}

load(file.path(PATH_RESULTS, file = "data.processed.RData"))

require(biostatUZH)
require(tidyverse)
require(meta)
require(metasens)
require(gridExtra)
require(xtable)
@

>>>>

load(file.path(PATH_RESULTS, "meta_analyses_summary_complete.RData"))
load(file.path(PATH_RESULTS, "meta.bin.RData"))
load(file.path(PATH_RESULTS, "meta.cont.RData"))
load(file.path(PATH_RESULTS, "meta.surv.RData"))
load(file.path(PATH_RESULTS, "data_used_for_analysis.RData"))

#Explanation of coding: 0 = (unsig, unsig), 1 = (unsig, sig), 2 = (sig, sig), 3 = (sig, unsig)
sig.level <- 0.05
meta.f <- meta.f %>% 
	mutate(
		pval.copas = 2*(1-pnorm(abs(est.copas/se.est.copas))),

		sig.fixef = ifelse(pval.fixef > sig.level, 0, 1),
		sig.ranef = ifelse(pval.ranef > sig.level, 0, 1),
		sig.reg.ranef = ifelse(pval.reg > sig.level, 0, 1),
		sig.copas = ifelse(pval.copas > sig.level, 0, 1),

		sig.change.ranef.reg = sig.change(sig.before = sig.ranef, sig.after =  sig.reg), #Function to register if significance of estimate has changed after correction
		sig.change.ranef.copas = sig.change(sig.before = sig.ranef, sig.after =  sig.copas),
		sig.change.fixef.reg = sig.change(sig.before = sig.fixef, sig.after =  sig.reg),
		sig.change.fixef.copas = sig.change(sig.before = sig.fixef, sig.after =  sig.copas))

sig.level <- 0.1
meta.f <- meta.f %>% 
	mutate(egger.test = ifelse(pval.peter < sig.level, 1, 0),
				 thompson.test = ifelse(pval.thompson < sig.level, 1, 0),
				 begg.test = ifelse(pval.begg < sig.level, 1, 0),

				 schwarzer.test = ifelse(pval.schwarzer < sig.level, 1, 0),
				 rucker.test = ifelse(pval.rucker < sig.level, 1, 0),
				 harbord.test = ifelse(pval.harbord < sig.level, 1, 0),
				 peter.test = ifelse(pval.peter < sig.level, 1, 0))

meta.bin <- meta.f %>% filter(outcome.type == "bin")
meta.cont <- meta.f %>% filter(outcome.type == "cont")
meta.surv <- meta.f %>% filter(outcome.type == "surv")

#Comparison of "original effect sizes":
effect.diff <- meta.f %>% mutate(
	est.fixef = ifelse(outcome.type == "bin", exp(est.fixef), est.fixef),
	est.ranef = ifelse(outcome.type == "bin", exp(est.ranef), est.ranef),
	est.copas = ifelse(outcome.type == "bin", exp(est.copas), est.copas),
	est.reg = ifelse(outcome.type == "bin", exp(est.reg.ranef), est.reg.ranef)
)

effect.diff <- effect.diff %>% mutate(
	est.fixef = ifelse(outcome.type == "bin", est.fixef - 1, est.fixef),
	est.ranef = ifelse(outcome.type == "bin", est.ranef - 1, est.ranef),
	est.copas = ifelse(outcome.type == "bin", est.copas - 1, est.copas),
	est.reg = ifelse(outcome.type == "bin", est.reg.ranef - 1, est.reg.ranef),
	fixef.ranef = ifelse(abs(est.ranef) > abs(est.fixef), "Reduction", "Amplification"),
	fixef.copas = ifelse(abs(est.fixef) > abs(est.copas), "Reduction", "Amplification"),
	fixef.reg = ifelse(abs(est.fixef) > abs(est.reg), "Reduction", "Amplification"),
	fixef.copas.s = ifelse(sign(est.fixef) == sign(est.copas), "Unchanged", "Reversed"),
	fixef.reg.s = ifelse(sign(est.fixef) == sign(est.reg), "Unchanged", "Reversed"))

@
